---
title: "stringr and regular expressions"
author: "Jiaying Wu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE)
```

This chapter will focus on the stringr package for string manipulation, which is part of the core tidyverse.

```{r}
library(tidyverse)
```

### 1. String basics

You can create strings with either single quotes or double quotes. Yor are recommended to always use `"`, unless you want to create a string that contains multiple `"`.

```{r}
string1 <- "This is a string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```

To include a literal single or double quote in a string you can use `\` to “escape” it:

```{r}
# or '"'
double_quote <- "\"" # or '"'
# or "'"
single_quote <- '\'' # or "'"
```

That means if you want to include a literal backslash, you’ll need to double it up: `"\\"`.

```{r}
x <- c("\"", "\\")
x
```

To see the raw contents of the string, use `writeLines()`:

```{r}
writeLines(x)
```

Multiple strings are often stored in a character vector, which you can create with `c()`:

```{r}
c("one", "two", "three")
```

#### 1.1 String length

`str_length()` tells you the number of characters in a string:

```{r}
str_length(c("a", "R for data science", NA))
```

The common `str_` prefix is particularly useful if you use RStudio, because typing `str_` will trigger autocomplete, allowing you to see all stringr functions:

```{r eval=FALSE}
str_
```

#### 1.2 Combining strings

To combine two or more strings, use `str_c()`:

```{r}
str_c("x", "y")
str_c("x", "y", "z")
```

Use the `sep` argument to control how they’re separated:

```{r}
str_c("x", "y", sep = ", ")
```

#### 1.3 Subsetting strings

You can extract parts of a string using `str_sub()`. `str_sub()` takes `start` and `end` arguments which give the (inclusive) position of the substring:

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
str_sub(x, -3, -1)
```

Note that `str_sub()` won’t fail if the string is too short: it will just return as much as possible:

```{r}
str_sub("a", 1, 5)
```

### 1.4 Locales

You can use `str_to_lower()` to change the text to lower case. You can also use `str_to_upper()` or `str_to_title()`.

```{r}
str_to_lower(x)
str_to_upper(c("a", "b"))
```

If you want robust behaviour across different computers, you may want to use str_sort() and str_order() which take an additional locale argument:

```{r}
x <- c("apple", "eggplant", "banana")
str_sort(x, locale = "en")  # English
str_sort(x, locale = "haw") # Hawaiian
```

### 2. Matching patterns with regular expressions

Regexps are a very terse language that allow you to describe patterns in strings. To learn regular expressions, we’ll use `str_view()` and `str_view_all()`. 

#### 2.1 Basic matches

The simplest patterns match exact strings:

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

The next step up in complexity is `.`, which matches any character (except a newline):

```{r}
str_view(x, ".a.")
```

Like strings, regexps use the backslash, `\`, to escape special behaviour. So to match an `.`, you need the regexp `\.`. Unfortunately this creates a problem. We use strings to represent regular expressions, and `\` is also used as an escape symbol in strings. So to create the regular expression `\.` we need the string `"\\."`.

```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

```{r}
# And this tells R to look for an explicit .
str_view(c("abc", "a.c", "bef"), "a\\.c")
```

In order to match a literal `\`,  you need to escape it, creating the regular expression `\\`. To create that regular expression, you need to use a string, which also needs to escape `\`. That means to match a literal `\` you need to write `"\\\\"`

```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

#### 2.2 Anchors

You can use:

- `^` to match the start of the string.

- `$` to match the end of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
str_view(x, "a$")
```

To force a regular expression to only match a complete string, anchor it with both `^` and `$`:

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

#### 2.3 Character classes and alternatives

You’ve already seen `.`, which matches any character apart from a newline. There are four other useful tools:

- `\d`: matches any digit.

- `\s`: matches any whitespace (e.g. space, tab, newline).

- `[abc]`: matches a, b, or c.

- `[^abc]`: matches anything except a, b, or c.

Remember, to create a regular expression containing `\d` or `\s`, you’ll need to escape the `\` for the string, so you’ll type `"\\d"` or `"\\s"`.

```{r}
# Look for a literal character that normally has special meaning in a regex
str_view(c("abc", "a.c", "a*c", "a c"), "a[.]c")
```

```{r}
str_view(c("abc", "a.c", "a*c", "a c"), ".[*]c")
```

```{r}
str_view(c("abc", "a.c", "a*c", "a c"), "a[ ]")
```

This works for most (but not all) regex metacharacters: `$` `.` `|` `?` `*` `+` `(` `)` `[` `{`. Unfortunately, a few characters have special meaning even inside a character class and must be handled with backslash escapes: `]` `\` `^` and `-`.

You can use `|` to pick between one or more alternative patterns.

```{r}
str_view(c("grey", "gray"), "gr(e|a)y")
```

#### 2.4 Repetition

The next step up in power involves controlling how many times a pattern matches:

- `?`: 0 or 1

- `+`: 1 or more

- `*`: 0 or more

```{r}
x <- "MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "C+")
```

```{r}
str_view(x, 'C[LX]+')
```

You can also specify the number of matches precisely:

- `{n}`: exactly n

- `{n,}`: n or more

- `{,m}`: at most m

- `{n,m}`: between n and m

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, "C{2,3}")
```

By default these matches are “greedy”: they will match the longest string possible. You can make them “lazy”, matching the shortest string possible by putting a `?` after them. 

```{r}
str_view(x, "C{2,3}?")
```

```{r}
str_view(x, 'C[LX]+?')
```

#### 2.5 Grouping and backreferences

Parentheses also create a numbered capturing group (number 1, 2 etc.). A **capturing group** stores the part of the string matched by the part of the regular expression inside the parentheses. You can refer to the same text as previously matched by a capturing group with **backreferences**, like `\1`, `\2` etc. 
 
For example, the following regular expression finds all fruits that have a repeated pair of letters.

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```


### Reference

Hadey, W. & Grolemund, G. (2017). R for Data Science. O’Reilly.

